Robustness of MinD oscillation in \emph{Escherichia coli} with diverse cell shapes

### 2014-09-19 paper/paper.tex
changes to intro of paper, changed verbiage to model from method throughout paper, and updated some data Tested everything on quipu.

Variations of the Huang 2003 model that stochasticly account for
molecular interaction and
diffusion~\cite{kerr2006division,fange2006noise} confirm the major
results obtained by Huang's model, and more successfully predict
experimentally observed oscillations in round cell
phenotypes~\cite{fange2006noise, huang2004min}. The successful
simulations of three pronged cells mentioned above are stochastic
simulations~\cite{varma2008min}. Biochemical models of broader scope
have also been used to study the MinD system and show consistent
results~\cite{arjunan2010new}.

\fixme{There are not any direct comparisons of the deterministic and
  stochastic methods that I can find other than the fange2006 paper
  cited above for the round cells.  Those showed that the
  deterministic sims are 'bistable' - one type of initial conditions
  (MinD starting on walls on one side) show oscillations, another type
  (half cyto is filled with 3/4 of MinD) don't.  Huang in 2004 paper
  cited above uses numerical determinisitc and finds that it can show
  oscillations in round cells as well, but right now I can't find it
  and the journal doesn't give me access to it.  predict oscillations
  in perfectly round cells unless noise was added to the simulation,
  but could in almost round ellipses.  It seems that once the
  stochastic methods began to be used, nobody really used the
  deterministic method any longer.  Not sure what to say about this in
  order to introduce what we do here properly.}

In this paper we apply a deterministic method and a stochastic method
of simulation and compare them against Mannik's experimental findings.

@@

We have performed both a basic finite analysis, numerical,
deterministic simulation that is spatially and temporally discrete,
and a stochastic analysis that is spatially discrete but continuous in
time.  Our stochastic simulation method follows the work of
Kraus~\cite{kraus1996crosstalk} which in turn follows a method
introduced by Gillespie~\cite{gillespie1977exact}.

@@

\begin{figure*}
  %\includegraphics[width=\textwidth]{../data/shape-p/3_00-0_50-0_00-0_00-15_00-exact/plots/image-plot}
  \includegraphics[width=\textwidth]{../data/shape-p/3_00-0_50-0_00-0_00-15_00-exact/plots/single-image-plot}
  \includegraphics[width=\textwidth]{../data/shape-p/3_00-0_50-0_00-0_00-15_00-full_array/plots/single-image-plot}
  \caption{Contour plot images of the concentration of each protein
    species in a natural pill-shaped bacterium at regular intervals in
    time (one second intervals), with darker regions indicating higher
    concentration. The upper plots shows results from the
    deterministic simulation and the lower shows results from the
    stochastic simulation.  The cells pictured are $4\micron$ in
    length, measured from end to end.  The order of frames is such
    that individual MinD proteins begin at the bottom of the plot (in
    the MinD:ATP state in the cytoplasm), and progress upward until
    they reach the MinE:MinD:ATP membrane-bound complex.  At that
    point, they will spontaneously dissociate into cytoplasmic MinE
    (the top row) and the starting state of cytoplasmic MinD:ADP.  The
    densities plotted are integrated along the axis orthogonal to the
    page, and the color scale is chosen separately for each species
    with black as the maximum value over space and time.  The
    stochastic densities are smeared out over space using a guassians
    approximation developed by Zhang et all~\cite{zhang2007gaussian}}.
  \label{image-p}
\end{figure*}

Huang \emph{et al}~\cite{huang2003dynamic} has performed a linear
stability analysis on the model which shows an upper limit on a steady
state solution of a $2\mu m$ wavelength.  Cells with dimensions longer
than this are unstable and show oscillatory behavoir in this
direction, while cells that are shorter relax into a motionless steady
state.  We have similarly performed a stability analysis in an
infinite slab with a thickness equal to that of our flattened cells'
($0.25\mu m$) and have found that an equivalent stability limit for
half wavelengths is $2.13\mu m$. As expected, when decreasing the
lengths and widths of our simulated flattened cells so that the
longest distance across the cell is less than this length, the cells
stop exhibiting any oscillalitory behavior.  The deterministic
simulations relax into a motionless state and the stochastic
simulations exhibit multiple random locations of small maximization.

@@

This simple cell shape is a good starting point for observing in
detail the dynamic interaction between the different stages of protein
that lead to their qualitatively distinct behavoir. Figure
\ref{image-p} shows the process in a series of frames taken from
simulation animations for both the deterministic and stochastic
simulations.  We have 'smeared out' the stochastic concentrations in a
manner meant to reproduce the images shown by diffraction limited
flourescence microscopy.  We do so using the two dimensional gaussian
approximation developed by Zhang et all~\cite{zhang2007gaussian}.  In
this approximation we use a numerical aperture value of 1.3, which is
the same as used by mannik experimentally, and a wavelength of
$650nm$.  Each frame is 2.5 seconds ahead of the last, and each image
shows the concentration of a given state of protein (of the five
described in the reaction model) summed over the coordinate orthogonal
to the page.  The color scale for each protein state is set according
to the maximum values observed.

Figure~\ref{image-p} begins about 300 seconds into the simulation and
shows a typical period of oscillation that in the pill shape repeates
itself with a high degree of regularity, for both the deterministic
and stochastic simulations.  At $t=0$ the cytoplasmic MinE is
concentrated in the upper portion of the cell and is diffusing
downward, where it will react with and stick to wall-bound MinD that
is concentrated on the membrane in the lower portion of the cell.
This process is in fact already well on its way, as seen by the
significant concentration of wall bound MinD:MinE which is creeping
down to the lower corner, removing the MinD from the membrane as it
goes.  The removed MinD (cytoplasmic MinD:ADP) diffuses freely for a
time before spontenously undergowing nucleotide exchange and shifting
back into the MinD:ATP that is able to attach to the membrane.
Diffusing a small distance away from the bottom of the cell will bring
the protein into the MinE collection, where reactions will prevent it
from collecting for any significant time on the membrane.  Diffusing
further towards the upper end, however, while the MinE has not yet
collected there, will allow MinD:ATP to accumulate with the other
MinD:ATP, which is seen at seconds 10 through 15 in
Figure~\ref{image-p}.  Eventually the MinE creeps all the way down to
the bottom of the cell where it slowly runs out of wall bound MinD to
react with, before diffusing back up and starting again on the upper
portion of the cell. This is seen in seconds 15 through 20 in the
figure.  \fixme{There is another paragraph after this in the latex
  that I commented out because I thought we might be spending too much
  time on the description}

@@

We test cells of radius $.5\micron$ and of lengths $2\micron$,
$3\micron$, $4\micron$, and $5\micron$ measured from the tip of one
end cap to the tip of the other \fixme{I'm running the 2,3,5 pills
  again to make sure that everything in this paragraph is true}. In
each case, we observe in the deterministic simulations a quick
establishement (within a few periods) of regular oscillatory behavoir
that lasts indefinitely.  This temporal regularity can be seen in
Figure~\ref{corr-pill}, which shows the temporal correlation of the
total MinD protein located in the the two polar sections of the cell.
In the stochastic simulations the pills exhibit coherence times of
over 7 periods and the determinisitic version cohere indefinately.

@@

While the stochastic results show periods that seem to have a high
degree of regularity as well as similarity to the deterministic
results, they do differ in the precise locations of the maxima.  While
the deterministic simulations show maxima that are every time
concentrated in the center of one of the cell ends, with a symmetric
smear of protein going out laterally to either side, the stochastic
simulations show maxima that, while always located somehwere in the
polar region of the cell, occur at different positions within this
region.  This is reflected in Figure~\ref{image-p}, which shows
periods that are the same and qualitative behavoir that is
approximately the same, but also that there is variation in the
details of the stochastic simulation maxima location.

@@

\begin{figure*}
  \centering
  \includegraphics[width=\textwidth]{../paper/plot-ave}
  \caption{We display here arrows depicting successive maxima in space
    and and time overlayed on a color plot of the total MinD density
    averaged over the same time period.  The simulation time covered
    for the flattened cells is \fixme{340} seconds, which is the same
    period of time depicted in the experimental data plots of Mannik
    \emph{et al.}~\cite{mannik2012robustness}.  For the wild type pill
    shape, we only cover \fixme{???} seconds, in order to provide a
    useful comparison due to its shorter oscillation period.  The top
    row shows plots published by Mannik of the MinD maxima behavoir
    and the bottom two rows show our simulations using the stochastic
    and deterministic algorithms, respectively.  We simulated
    approximations to the two shapes observed by Mannik, which we call
    \emph{shape A} and \emph{shape B}.  In addition, we studied two
    flattened stadium shapes which we call \emph{stadium A} and
    \emph{stadium B} corresponding in aspect ratio and thickness to
    the two experimental shapes.  The spatial length scale of all the
    figures shown was identical. Each of the flattened cell shapes
    uses the same color scale for the number of proteins per unit
    area.  Finally, we display the natural pill shape, which was also
    featured in Figs.~\ref{image-p} and~\ref{corr-pill}, with a
    different color scale to reflect the thicker cell containing more
    proteins per cross-sectional area.  }
  \label{randst-plot-ave}
\end{figure*}

@@

As describe in Section~\ref{sec:model-method-shapes}, we will focus on
simulation of four illustrative flattened cell shapes: two shapes
created to replicate those published by
Mannik~\cite{mannik2012robustness} (shape~\emph{A} and
shape~\emph{B}), and for comparison two symmetrical stadium shapes
that have the identical aspect ratio, thickness and volume (stadium
\emph{A} and stadium \emph{B}).  These can be seen in
Figure~\ref{randst-plot-ave}.  These stadium shapes enable us to
distinguish between the effect of cellular shape irregularity and
assymetry, and that of overall size and aspect ratio.  We note that
shape \emph{B} is more irregular than shape \emph{A}, and in
particular features a prominant region of concavity on its left-hand
side.  For each of these shapes (in addition to the wild type pill
shape discussed previously) we have simulated in excess of \fixme{???}
seconds of evolution of the MinD system using the deterministic method
of Huang's model~\cite{huang2003dynamic}, as well as the stochastic
method.

@@

In every case, including the wild type pill shape, we see irregularity
in the location of the maxima when using the stochastic approach.  The
deterministic method shows uniformly bipolar oscillation, with the
exception of shape \emph{B}, which develops weak intermediate maxima
on the right-hand edge of the cell.  From these results, we conclude
that the deterministic method is inadequate to explain experimental
observations of the locations of density maxima of the MinD protein.

@@

The predictions of the stochastic as seen in
Fig.~\ref{randst-plot-ave} are roughly similar to the experimental
results in terms of irregularity of the locations of maxima, although
there are locations on the edges of the cells where experiment shows
maxima occuring that we never see in our simulations, suggesting that
the model does not precisely reflect the experiment.  We also note
that the stadium shapes in Fig.~\ref{randst-plot-ave} appear
qualitatively similar to the irregular shapes in terms of the
locations of their maxima.

However, Fig.~\ref{randst-plot-ave} leaves unclear the importance of
irregularity versus overall thickness and aspect ratio in developing
an understanding of the origin of the experimentally observed
behavior.  Although in the movies the stadiums appear to display a
more regular bipolar oscillation, the arrow plots do not provide a
convincing demonstration.  To illustrate the difference in behavior
between shape \emph{A} and stadium \emph{A}, we return in
Fig.~\ref{fig:corr-pancake-A} to the correlation function between the
number of proteins in a segment at the top of the cell and the number
of proteins in a segment at the bottom of the cell.  We computed these
correlation functions using over \fixme{???} seconds of simulation.
We found with the stochastic method that the experimental shape
\emph{A} has a coherence time of under 4 periods, while the
corresponding stadium \emph{A} has a coherence time of 10 periods.  In
each case, the deterministic method is perfectly periodic. This
confirms that using the stochastic method the stadium shape does
indeed have a far more regular bipolar oscillation than the irregular
shape \emph{A}.  When we compare shape \emph{B} and stadium \emph{B},
we see a similar effect, although both coherence times are somewhat
smaller (2.7 and 6.2, respectively).  From these results we conclude
that the irregularity seen experimentally is indeed exacerbated by the
irregular shapes taken on by the cells forced into a thin slit.

\section{Conclusion}
We have simulated flattened cell shapes similar to those observed
experimentally by Mannik \emph{et al.}.  We find that the
deterministic method~\fixme{cite} predicts behavior that completely
contradicts what is seen experimentally: it predicts stable and
periodic bipolar oscillation, with a minor intermediate maximum
appearing in shape \emph{B}.  Thus we conclude that this method is
inconsistent with experiment, and should not be relied upon for
predictions of the behavior of the MinD system.

As found in previous studies~\fixme{cite}, we observe that the
stochastic method predicts behavior in wild type cells that is
consistent both with experiment and with the deterministic method.  In
contrast, the stochastic method~\fixme{cite} predicts spatially
irregular formation of maxima for all the flattened cells that we
studied.  In order to better understand the irregular oscillations
observed in experiment, we simulated flattened ``stadium'' shapes in
order to distinguish between the effect of flattening and the effect
of irregular shape.  We found that although the stadium cells
exhibited somewhat more irregular locations of maxima than the
wild-type cells, their oscillation was far more periodic and bipolar
than the irregular cell shapes observed experimentally.  Thus we
confirm that the disruption of the MinD oscillations seen by Mannik
\emph{et al.} is largely due to the irregularity of cell shapes
observed, as opposed to the size or aspect ratio.

>>>>

Variations of the Huang 2003 model that stochasticly account for
molecular interaction and diffusion~\cite{fange2006noise} and
monte-carlo simulations that implement this stochastic version of
Huang's mean field reaction rates~\cite{kerr2006division} broadly
confirm the major results obtained by Huang's model.  However, they
more successfully predict experimentally observed oscillations in
round cell phenotypes~\cite{fange2006noise, huang2004min} and
variations in other
quantities~\cite{kruse2007experimentalist}\fixme{If going to use this
  sentence should list these and reference properly?}. Both
deterministic and stochastic models are widely used throughout systems
biology~\cite{lawson2013spatial,robb2014stochastic,oguz2014stochastic,fu2013deterministic,rudiger2014stochastic},
and it's important to better understand their limitations.  Manniks'
experiments allow for a new distinct opportunity to test these two
approaches further.
%% Biochemical models of broader scope have also been
%% used to study the MinD system and show consistent
%% results~\cite{arjunan2010new}.

%% \fixme{There are not any direct comparisons of the deterministic and
%%   stochastic models that I can find other than the fange2006 paper
%%   cited above for the round cells.  Those showed that the
%%   deterministic sims are 'bistable' - one type of initial conditions
%%   (MinD starting on walls on one side) show oscillations, another type
%%   (half cyto is filled with 3/4 of MinD) don't.  Huang in 2004 paper
%%   cited above uses numerical determinisitc and finds that it can show
%%   oscillations in round cells as well, but right now I can't find it
%%   and the journal doesn't give me access to it.  predict oscillations
%%   in perfectly round cells unless noise was added to the simulation,
%%   but could in almost round ellipses.  It seems that once the
%%   stochastic models began to be used, nobody really used the
%%   deterministic model any longer.  Not sure what to say about this in
%%   order to introduce what we do here properly.}

In this paper we apply a deterministic model and a stochastic model
of simulation and compare them against Mannik's experimental findings.

@@

We have performed both a basic finite analysis, numerical,
deterministic model simulation that is spatially and temporally
discrete, and a stochastic analysis that is spatially discrete but
continuous in time.  Our stochastic model follows the work of
Kraus~\cite{kraus1996crosstalk} which in turn follows a method
introduced by Gillespie~\cite{gillespie1977exact}.

@@

\begin{figure*}
  %\includegraphics[width=\textwidth]{../data/shape-p/3_00-0_50-0_00-0_00-15_00-exact/plots/image-plot}
  \includegraphics[width=\textwidth]{../data/shape-p/3_00-0_50-0_00-0_00-15_00-exact/plots/single-image-plot}
  \includegraphics[width=\textwidth]{../data/shape-p/3_00-0_50-0_00-0_00-15_00-full_array/plots/single-image-plot}
  \caption{Contour plot images of the concentration of each protein
    species in a natural pill-shaped bacterium at regular intervals in
    time (one second intervals), with darker regions indicating higher
    concentration. The upper plots shows results from the
    deterministic model and the lower shows results from the
    stochastic model.  The cells pictured are $4\micron$ in
    length, measured from end to end.  The order of frames is such
    that individual MinD proteins begin at the bottom of the plot (in
    the MinD:ATP state in the cytoplasm), and progress upward until
    they reach the MinE:MinD:ATP membrane-bound complex.  At that
    point, they will spontaneously dissociate into cytoplasmic MinE
    (the top row) and the starting state of cytoplasmic MinD:ADP.  The
    densities plotted are integrated along the axis orthogonal to the
    page, and the color scale is chosen separately for each species
    with black as the maximum value over space and time.  The
    stochastic densities are smeared out over space using a guassians
    approximation developed by Zhang et all~\cite{zhang2007gaussian}}.
  \label{image-p}
\end{figure*}

Huang \emph{et al}~\cite{huang2003dynamic} has performed a linear
stability analysis on the model which shows an upper limit on a steady
state solution of a $2\mu m$ wavelength.  Cells with dimensions longer
than this are unstable and show oscillatory behavoir in this
direction, while cells that are shorter relax into a motionless steady
state.  We have similarly performed a stability analysis in an
infinite slab with a thickness equal to that of our flattened cells'
($0.25\mu m$) and have found that an equivalent stability limit for
half wavelengths is $2.13\mu m$. As expected, when decreasing the
lengths and widths of our simulated flattened cells so that the
longest distance across the cell is less than this length, the cells
stop exhibiting any oscillalitory behavior.  The deterministic model
relaxes into a motionless state and the stochastic model exhibits
multiple random locations of small maximization.

@@

This simple cell shape is a good starting point for observing in
detail the dynamic interaction between the different stages of protein
that lead to their qualitatively distinct behavoir. Figure
\ref{image-p} shows the process in a series of frames taken from
simulation animations for both the deterministic and stochastic
models.  We have 'smeared out' the stochastic concentrations in a
manner meant to reproduce the images shown by diffraction limited
flourescence microscopy.  We do so using the two dimensional gaussian
approximation developed by Zhang et all~\cite{zhang2007gaussian}.  In
this approximation we use a numerical aperture value of 1.3, which is
the same as used by mannik experimentally, and a wavelength of
$650nm$.  Each frame is 2.5 seconds ahead of the last, and each image
shows the concentration of a given state of protein (of the five
described in the reaction model) summed over the coordinate orthogonal
to the page.  The color scale for each protein state is set according
to the maximum values observed.

Figure~\ref{image-p} begins about 300 seconds into the simulation and
shows a typical period of oscillation that in the pill shape repeates
itself with a high degree of regularity, for both the deterministic
and stochastic models.  At $t=0$ the cytoplasmic MinE is
concentrated in the upper portion of the cell and is diffusing
downward, where it will react with and stick to wall-bound MinD that
is concentrated on the membrane in the lower portion of the cell.
This process is in fact already well on its way, as seen by the
significant concentration of wall bound MinD:MinE which is creeping
down to the lower corner, removing the MinD from the membrane as it
goes.  The removed MinD (cytoplasmic MinD:ADP) diffuses freely for a
time before spontenously undergowing nucleotide exchange and shifting
back into the MinD:ATP that is able to attach to the membrane.
Diffusing a small distance away from the bottom of the cell will bring
the protein into the MinE collection, where reactions will prevent it
from collecting for any significant time on the membrane.  Diffusing
further towards the upper end, however, while the MinE has not yet
collected there, will allow MinD:ATP to accumulate with the other
MinD:ATP, which is seen at seconds 10 through 15 in
Figure~\ref{image-p}.  Eventually the MinE creeps all the way down to
the bottom of the cell where it slowly runs out of wall bound MinD to
react with, before diffusing back up and starting again on the upper
portion of the cell. This is seen in seconds 15 through 20 in the
figure.  \fixme{There is another paragraph after this in the latex
  that I commented out because I thought we might be spending too much
  time on the description}

@@

We test cells of radius $.5\micron$ and of lengths $2\micron$,
$3\micron$, $4\micron$, and $5\micron$ measured from the tip of one
end cap to the tip of the other \fixme{I'm running the 2,3,5 pills
  again to make sure that everything in this paragraph is true}. In
each case, we observe in the deterministic model a quick
establishement (within a few periods) of regular oscillatory behavoir
that lasts indefinitely.  This temporal regularity can be seen in
Figure~\ref{corr-pill}, which shows the temporal correlation of the
total MinD protein located in the the two polar sections of the cell.
In the case of the stochastic model the pills exhibit coherence times
of over 7 periods and in the case of the determinisitic model cohere
indefinately.

@@

While the stochastic results show periods that seem to have a high
degree of regularity as well as similarity to the deterministic
results, they do differ in the precise locations of the maxima.  While
the deterministic model shows maxima that are every time
concentrated in the center of one of the cell ends, with a symmetric
smear of protein going out laterally to either side, the stochastic
model shows maxima that, while always located somehwere in the polar
region of the cell, occur at different positions within this region.
This is reflected in Figure~\ref{image-p}, which shows periods that
are the same and qualitative behavoir that is approximately the same,
but also that there is variation in the details of the stochastic
model maxima location.

@@

\begin{figure*}
  \centering
  \includegraphics[width=\textwidth]{../paper/plot-ave}
  \caption{We display here arrows depicting successive maxima in space
    and and time overlayed on a color plot of the total MinD density
    averaged over the same time period.  The simulation time covered
    for the flattened cells is \fixme{340} seconds, which is the same
    period of time depicted in the experimental data plots of Mannik
    \emph{et al.}~\cite{mannik2012robustness}.  For the wild type pill
    shape, we only cover \fixme{???} seconds, in order to provide a
    useful comparison due to its shorter oscillation period.  The top
    row shows plots published by Mannik of the MinD maxima behavoir
    and the bottom two rows show our simulations using the stochastic
    and deterministic models, respectively.  We simulated
    approximations to the two shapes observed by Mannik, which we call
    \emph{shape A} and \emph{shape B}.  In addition, we studied two
    flattened stadium shapes which we call \emph{stadium A} and
    \emph{stadium B} corresponding in aspect ratio and thickness to
    the two experimental shapes.  The spatial length scale of all the
    figures shown was identical. Each of the flattened cell shapes
    uses the same color scale for the number of proteins per unit
    area.  Finally, we display the natural pill shape, which was also
    featured in Figs.~\ref{image-p} and~\ref{corr-pill}, with a
    different color scale to reflect the thicker cell containing more
    proteins per cross-sectional area.  }
  \label{randst-plot-ave}
\end{figure*}

@@

As describe in Section~\ref{sec:model-method-shapes}, we will focus on
simulation of four illustrative flattened cell shapes: two shapes
created to replicate those published by
Mannik~\cite{mannik2012robustness} (shape~\emph{A} and
shape~\emph{B}), and for comparison two symmetrical stadium shapes
that have the identical aspect ratio, thickness and volume (stadium
\emph{A} and stadium \emph{B}).  These can be seen in
Figure~\ref{randst-plot-ave}.  These stadium shapes enable us to
distinguish between the effect of cellular shape irregularity and
assymetry, and that of overall size and aspect ratio.  We note that
shape \emph{B} is more irregular than shape \emph{A}, and in
particular features a prominant region of concavity on its left-hand
side.  For each of these shapes (in addition to the wild type pill
shape discussed previously) we have simulated in excess of \fixme{???}
seconds of evolution of the MinD system using the deterministic
version of Huang's model~\cite{huang2003dynamic}, as well as the
stochastic model.

@@

In every case, including the wild type pill shape, we see irregularity
in the location of the maxima when using the stochastic model.  The
deterministic model shows uniformly bipolar oscillation, with the
exception of shape \emph{B}, which develops weak intermediate maxima
on the right-hand edge of the cell.  From these results, we conclude
that the deterministic model is inadequate to explain experimental
observations of the locations of density maxima of the MinD protein.

@@

The predictions of the stochastic model as seen in
Fig.~\ref{randst-plot-ave} are roughly similar to the experimental
results in terms of irregularity of the locations of maxima, although
there are locations on the edges of the cells where experiment shows
maxima occuring that we never see in our simulations, suggesting that
the model does not precisely reflect the experiment.  We also note
that the stadium shapes in Fig.~\ref{randst-plot-ave} appear
qualitatively similar to the irregular shapes in terms of the
locations of their maxima.

However, Fig.~\ref{randst-plot-ave} leaves unclear the importance of
irregularity versus overall thickness and aspect ratio in developing
an understanding of the origin of the experimentally observed
behavior.  Although in the movies the stadiums appear to display a
more regular bipolar oscillation, the arrow plots do not provide a
convincing demonstration.  To illustrate the difference in behavior
between shape \emph{A} and stadium \emph{A}, we return in
Fig.~\ref{fig:corr-pancake-A} to the correlation function between the
number of proteins in a segment at the top of the cell and the number
of proteins in a segment at the bottom of the cell.  We computed these
correlation functions using over \fixme{???} seconds of simulation.
We found with the stochastic model that the experimental shape
\emph{A} has a coherence time of under 4 periods, while the
corresponding stadium \emph{A} has a coherence time of 10 periods.  In
each case, the deterministic model is perfectly periodic. This
confirms that with the stochastic model the stadium shape does indeed
have a far more regular bipolar oscillation than the irregular shape
\emph{A}.  When we compare shape \emph{B} and stadium \emph{B}, we see
a similar effect, although both coherence times are somewhat smaller
(2.7 and 6.2, respectively).  From these results we conclude that the
irregularity seen experimentally is indeed exacerbated by the
irregular shapes taken on by the cells forced into a thin slit.

\section{Conclusion}
We have simulated flattened cell shapes similar to those observed
experimentally by Mannik \emph{et al.}.  We find that the
deterministic model~\fixme{cite} predicts behavior that completely
contradicts what is seen experimentally: it predicts stable and
periodic bipolar oscillation, with a minor intermediate maximum
appearing in shape \emph{B}.  Thus we conclude that this method is
inconsistent with experiment, and should not be relied upon for
predictions of the behavior of the MinD system.

As found in previous studies~\fixme{cite}, we observe that the
stochastic model predicts behavior in wild type cells that is
consistent both with experiment and with the deterministic model.  In
contrast, the stochastic model~\fixme{cite} predicts spatially
irregular formation of maxima for all the flattened cells that we
studied.  In order to better understand the irregular oscillations
observed in experiment, we simulated flattened ``stadium'' shapes in
order to distinguish between the effect of flattening and the effect
of irregular shape.  We found that although the stadium cells
exhibited somewhat more irregular locations of maxima than the
wild-type cells, their oscillation was far more periodic and bipolar
than the irregular cell shapes observed experimentally.  Thus we
confirm that the disruption of the MinD oscillations seen by Mannik
\emph{et al.} is largely due to the irregularity of cell shapes
observed, as opposed to the size or aspect ratio.
